Extraction once again

### 2021-06-30 paper.tex
Add a bit more structure to the paper.
        NOTE there is no much text in there.  Mostly thoughts
             and parts of the old paper.

\begin{document}
\section{Introduction}
Right now we have the following plan about the structure of the GPCE paper.
\begin{enumerate}
    \item Goal: implementing DSL with dependent type system 
    \item Problem: current two approaches are either doing a deep embedding or
        implementing a typechecker from scratch
    \item Reflection to the rescue: we can use reflection for implementing
        custom compilers 
    \item Our approach by example: demonstrate that this actually works in Agda
        (ps example)
    \item Extracting definitions by pattern matching: matching on the stack
    \item Using normalization to get rid of meta-functions (by using this
        technique, you get macros for free! / or use rewrite rules example:
        exchange o exchange = id)
    \item General technology for writing extractors
    \item This actually scales to realistic and useful examples!
\end{enumerate}

@@

>>>>

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}

\citestyle{acmauthoryear}   %% For author/year citations

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

% Add definitions for abbreviations like e.g.; i.e; etc. with
% correct spacing depending on the parameters.  In this case
% `all' exposes all the definitions of the package, and `british'
% makes sure that there is no comma after e.g. or i.e.
\usepackage[all,british]{foreign}

\usepackage{mathpartir}

\usepackage{todonotes}

\usepackage{varwidth}

\usepackage{agda}
\usepackage{newunicodechar}
\usepackage[final]{listings}

\newunicodechar{∷}{::}
\newunicodechar{→}{\ensuremath{\to}}
\newunicodechar{ω}{\ensuremath{\omega}}
\newunicodechar{⊎}{\ensuremath{\uplus}}
\newunicodechar{≔}{\ensuremath{\coloneqq}}
\newunicodechar{∎}{\ensuremath{\blacksquare}}
\newunicodechar{≟}{\ensuremath{\stackrel{?}{=}}}
\newunicodechar{ᵣ}{\ensuremath{_r}}
\newunicodechar{ₗ}{\ensuremath{_l}}
\newunicodechar{▴}{\ensuremath{\blacktriangle}}
\newunicodechar{▾}{\ensuremath{\blacktriangledown}}
\newunicodechar{▹}{\ensuremath{\triangleright}}
\newunicodechar{∸}{\ensuremath{\dot{-}}}
\newunicodechar{⊛}{\ensuremath{\circledast}}
\newunicodechar{ᵀ}{\ensuremath{{}^{\text{T}}}}
\newunicodechar{⌿}{\ensuremath{\not-}}

% Some shortcut commands for agda symbols
\newcommand{\AD}[1]{\AgdaDatatype{#1}}
\newcommand{\AC}[1]{\AgdaInductiveConstructor{#1}}
\newcommand{\AF}[1]{\AgdaFunction{#1}}
\newcommand{\AB}[1]{\AgdaBound{#1}}
\newcommand{\AK}[1]{\AgdaKeyword{#1}}
\newcommand{\AR}[1]{\AgdaField{#1}}
\newcommand{\AM}[1]{\AgdaModule{#1}}
\newcommand{\AN}[1]{\AgdaNumber{#1}}
\newcommand{\AS}[1]{\AgdaString{#1}}

\newcommand\codeblock[1]{%
  %{\fbox{\begin{varwidth}{0.9\textwidth}#1\end{varwidth}}}
  {\begin{varwidth}{0.9\textwidth}#1\end{varwidth}}
}

\begin{document}

%% Title information
\title[]{Extraction once again}

%% Author with single affiliation.
\author[A. {\v{S}}inkarovs]{Artjoms {\v{S}}inkarovs}
\orcid{0000-0003-3292-2985}
%\authornote{with author1 note}          %% \authornote is optional;
                                         %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{%
  \institution{Heriot-Watt University}   %% \institution is required
  \city{Edinburgh}
  \state{Scotland}
  \postcode{EH14 4AS}
  \country{UK}}
\email{a.sinkarovs@hw.ac.uk}             %% \email is recommended

%% Author with single affiliation.
\author[J. Cockx]{Jesper Cockx}
%\authornote{with author1 note}          %% \authornote is optional;
                                         %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{%
  \institution{TU Delft}
  \city{Delft}
  \postcode{2628 XE}
  \country{Netherlands}}
\email{j.g.h.cockx@tudelft.nl}           %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Dependently-typed host languages empower users to verify a wide
  range of properties of embedded languages and
  programs written in them.  Designers of such embedded languages
  are faced with a difficult choice between using a shallow or a deep
  embedding.  The former is easier to use because the entire
  infrastructure of the host langauge is immediately
  available. Meanwhile, the latter gives full access to the
  structure of embedded programs, but is difficult to use in practice,
  especially when the embedded language is itself dependently typed.

  The main insight presented in this paper is that the choice between
  shallow and deep embedding can be eliminated by working in a host
  language with reflection capabilities: we start from a shallow
  embedding that can use all libraries and tools of the host language,
  and later use reflection to expose the deep structure of the
  embedded programs.
  %
  Concretely, we apply this technique to embed three
  programming languages --- Kaleidoscope, SaC, and (a subset of) APL ---
  into the dependently typed theorem prover Agda, using dependent
  types to statically enforce several properties of interest.  We then
  use Agda's reflection capabilities to extract the embedded programs
  back into the original language, so that the existing toolchain can
  be leveraged.
  %
  In this process, statically verified properties of the host language
  are mapped onto runtime checks in the target language, allowing
  extracted programs to interact safely with existing code.
  %
  Finally, we demonstrate the feasibility of our approach with the
  implementation and extraction of a convolutional neural network in
  our embedding of APL.\@

%  Most practical programming languages of today prioriterise
%  performance and usability over safety.  As a result, many
%  useful correctness invariants cannot be expressed in the given
%  language, typically due to weakness of the chosen type system.
%  We are looking for non-invasive verification: how to
%  guarantee a property about the given program without any modifications
%  to the language toolchain.
%
%  We study an approach when a programming language is shallowly
%  embedded into a theorem prover, and dependent types are used
%  to encode the properties of interest.   Bringing the embedded
%  langauge and the properties back into the original language
%  is the main focus of this work.  We propose a generic extraction
%  mechanism that we implement in Agda that is based on the concept
%  of reflection.  Making it possible to translate any shallowly
%  embedded language in the target backend without the necessity
%  to modify the theorem prover.
%  We demonstrate our framework in action by encoding three languages:
%  Kaleidoskope, SaC and a subset of APL, culminating in the implementation
%  and extraction of a convolutional neural network.
\end{abstract}
\maketitle

\section{Introduction}
Right now we have the following plan about the structure of the GPCE paper.
\begin{enumerate}
    \item Goal: implementing DSL with dependent type system 
    \item Problem: current two approaches are either doing a deep embedding or
        implementing a typechecker from scratch
    \item Reflection to the rescue: we can use reflection for implementing
        custom compilers 
    \item Our approach by example: demonstrate that this actually works in Agda
        (ps example)
    \item Extracting definitions by pattern matching: matching on the stack
    \item Using normalization to get rid of meta-functions (by using this
        technique, you get macros for free! / or use rewrite rules example:
        exchange o exchange = id)
    \item General technology for writing extractors
    \item This actually scales to realistic and useful examples!
\end{enumerate}

@@

% Agda Basics and Extraction
\input{background}

% PostScript Language and its embedding
\section{PostScript and its embedding in Agda}
Demonstrate a few examples.

\section{Extraction}
\subsection{Framework Overview}
Explain very briefly that there is a common extraction part that we can (and do)
abstract away; and there is a language part that we are focusing on.  This can be
moved towards the end.

\subsection{PostScript Programs Repesentation}
Explain the AST and show encoding of an example.

\subsection{Normalisation} \label{sec:normalisation}
It is useful for inlining functions, and creating macros that can use arbitrary expressions
as long as their applications evaluate to postscript operators.
This is not essential and can be moved towards the end.

%\subsubsection{Telescopes}
\subsection{Controlling Reduction} \label{sec:controlling-reduction}
This is essential, as we need to avoid inlining of the base operators, otherwsie
we won't be able to extact our pograms one-to-one.  Explain the mechanism.  Say that we have
added this to Agda?

\subsection{\label{sec:maptypes}Analysing Types}
Explain that all we need to do is to find the position of the stack argument, as well as
verify that the return type is also a stack argument.  In between we can have arbitrary
number of computationally irrelevant arguments.

\subsection{Pattern Matching}
The natural way to express functions in Agda is by means of pattern-matching, so it makes sense
to support it.  All we allow in patternmatching is to ``look'' at the stack argument and the
individual elements.  But we are not allowed to reshuffle the elements, we have to pass the
pattern exactly as it was matched (in some sense it is read-only pattern matching).

\subsection{Translating terms} \label{sec:translating-terms}
The core of the extraction is how do we translate terms.  Also this is the simplest part in
case of the PostScript example.

\subsection{Example}
Explain the entire extaction process for the given example?

Shall we talk about termination and how do we deal with that?

\subsection{\label{sec:rewriting}Rewriting}
We can mention that rewriting is super useful feature that can be used as an optimisation
prior to extraction.

\section{Scaling to Bigger Langauges}
Not sure whether this deserves a separate section, but we need to explain
that the proposed technology perfectly scales to the level of bigger (more complex)
languages.

\section{Related Work}

\section{Conclusios and Future Work}

%% Bibliography
\bibliography{paper}
